Title

### 2020-06-30 paper2/paper2.tex
Added introduction and feature extraction

\section{Introduction}
@Boris

\section{Feature Extraction}
@Boris

\section{Neural Network}
The created dataset will be used in the further course of this paper to build a neural network. With this it should be possible to detect anomalies in network traffic (e. g. DoS attacks) but also harmless data packets. This is a continuation of  \cite{max1}, in which an attempt was made to create a neural network for intrusion detection on the basis of the publicly available CSE-CIC-IDS2018 dataset  \cite{max2}. The knowledge gained at that time is used, the source codes are reused accordingly and slightly adapted to the new dataset. In principle, the performance of neural networks over different datasets is to be checked. The following scenarios will be considered:
\smallskip

@@

\section{Conclusion}
In summary, the performance of neural networks in terms of intrusion detection over different datasets is rather mixed. If these machine learning algorithms are tested with the same dataset with which they were trained, the result is promising. Using another dataset for testing results in a poor classification performance. However, there is lots of potential for future research with neuronal networks in terms of intrusion detection. It seems useful to analyse in detail why the mentioned difficulties occur in the classification and which factors are relevant for the neural network. An intensive examination of the finding of optimal hyperparameters would also be desirable in the course of future work. Furthermore, it would also make sense to simplify the way classification is done. Instead, a simple classification like network traffic is “benign” or “not benign” would be conceivable. Moreover, it should be checked whether a better performance can be achieved with other machine learning algorithms.

\begin{thebibliography}{00}
\bibitem{max1} F. Ahrens, C. Cat and R. Freingruber, "A Neural-Network Approach for an Intrusion Detection System", FH Technikum, Vienna, 2019.
\bibitem{max2} I. Sharafaldin, A. H. Lashkari and A. A. Ghorbani, "Toward Generating a New Intrusion Detection Dataset and Intrusion Traffic Characterization", in 4th International Conference on Information Systems Security and Privacy (ICISSP), Portugal, 2018.
\bibitem{max3} D. Meyer, "Performance Assessment \& Tuning - Presentation slides", FH Technikum, Vienna, 2018.
\bibitem{max4} J. Brownlee, "How to Calculate Correlation Between Variables in Python", 3 June 2020. [Online]. Available: https://machinelearningmastery.com/how-to-use-correlation-to-understand-the-relationship-between-variables/. [Accessed 27 June 2020].
\bibitem{max5} S. Norena, "Finding Correlation Between Many Variables (Multidimensional Dataset) with Python", 26 April 2018. [Online]. Available: https://medium.com/@sebastiannorena/finding-correlation-between-many-variables-multidimensional-dataset-with-python-5deb3f39ffb3. [Accessed 27 June 2020].
\end{thebibliography}

>>>>

\section{Introduction}
The main goal of this paper is the comparison of a new dataset for intrusion prevention systems, which was created using automatically generated network traffic \cite{paper1}, with the CSE-CIC-IDS-2018 dataset \cite{max2}, using both a newly trained neural network and a neural network previously trained using the CSE-CIC-IDS-2018 dataset \cite{max1}. Because raw traffic network data cannot be used for training and classification using a neural network, meaningful features have to be extracted first. This process is described in Section \ref{extraction}. The extracted features can then be used for training and classification using multiple neural networks. This task is discussed and its results are shown in Section \ref{nn}. Finally, a conclusion based on these results is given in Section \ref{conclusion}.

\section{Feature Extraction}\label{extraction}
Before the neural network can be trained and used for classification, features have to be extracted from the raw network traffic data which was recorded in a previous step \cite{paper1}. For this task, the open-source Java software CICFlowMeter \cite{cicflowmeter} was chosen. The CICFlowMeter software was also used to create the CSE-CIC-IDS-2018 dataset \cite{max2}. This was a major reason for using it in this paper, as one of its goals is comparing the newly created dataset to the CSE-CIC-IDS2018 dataset. Some adaptations had to be made to the software, however, before it could be used to extract the desired features and also the labels from the recorded network traffic. Subsection \ref{extraction_using_cicflowmeter} describes how the CICFlowMeter in general extracts features from recorded network traffic. Subsection \ref{adaptation} then outlines the adaptations which had to be made.

\subsection{Extraction of flow-based features using CICFlowMeter}\label{extraction_using_cicflowmeter}
A common approach for extracting features from network traffic is to group packets into flows, and then calculate statistical features based on these flows. Because the CICFlowMeter software\cite{cicflowmeter} has been used used to fulfill that task for the creation of various datasets \cite{max2,sopuru,lashkari} and also during the creation of this paper, the following explanation is describing the functionality of that particular software. However, there is also a number of applications using various other software, also based on the idea of network-flow generation \cite{wang,moustafa,celik,waizumi}.

For each recorded packet read from a file in the \verb|pcap| format, the \verb|FlowGenerator| first checks if there is already a flow with that source IP address, destination IP address, source port, destination port and protocol (\verb|TCP| or \verb|UDP|). When this is the case, the packet information is added to that flow. Because CICFlowMeter is actually a bi-flow generator, the IP address and port of the source and destination may also be swapped, in which case the packet is regarded as a backwards packet. When no matching flow is found, a new flow is created (the first packet of each flow determines the forward and backward directions). A flow ends when a FIN packet is sent (TCP) or a configurable timeout occurs (UDP and TCP, 120 seconds by default), and the state of a flow changes from active to idle after a shorter timeout (default: five seconds). A flow is also divided into subflows whenever more than one second elapsed between two packets. The \verb|FlowGenerator| aggregates the information of each packet and calculates various statistical features. In addition to the features that define a flow (flow ID, IP address, port and protocol), CICFlowMeter supports a wide range of statistical features, including simple packet counts, totals, extrema (max, min), mean and standard deviation (Std). Some features, such as the inter-arrival-time (IAT, i.e. the time elapsed between two packets), are time-based, others depend on packet size, certain TCP flags and bulk (Blk) data flows. As the flows are bidirectional, the features can also distinguish between forward (Fwd) and backward (Bwd) directions. The full list of features supported by CICFlowMeter is given in Table \ref{table:cicflowmeter_features}, although the software is open-source and can therefore easily be extended to support currently disregarded features.

\begin{table}[htbp]
\caption{List of features supported by CICFlowMeter}
\begin{tabular}{| l l l |}
\hline
Flow ID & Src IP & Src Port \\
Dst IP & Dst Port & Protocol \\
Timestamp & Flow Duration & Tot Fwd Pkts \\
Tot Bwd Pkts & TotLen Fwd Pkts & TotLen Bwd Pkts \\
Fwd Pkt Len Max & Fwd Pkt Len Min & Fwd Pkt Len Mean \\
Fwd Pkt Len Std & Bwd Pkt Len Max & Bwd Pkt Len Mi \\
Bwd Pkt Len Mean & Bwd Pkt Len Std & Flow Bytes/s \\
Flow Pkts/s & Flow IAT Mean & Flow IAT Std \\
Flow IAT Max & Flow IAT Min & Fwd IAT Tot \\
Fwd IAT Mean & Fwd IAT Std & Fwd IAT Max \\
Fwd IAT Min & Bwd IAT Tot & Bwd IAT Mean \\
Bwd IAT Std & Bwd IAT Max & Bwd IAT Min \\
Fwd PSH Flags & Bwd PSH Flags & Fwd URG Flags \\
Bwd URG Flags & Fwd Header Len & Bwd Header Len \\
Fwd Pkts/s & Bwd Pkts/s & Pkt Len Min \\
Pkt Len Max & Pkt Len Mean & Pkt Len Std \\
Pkt Len Var & FIN Flag Cnt & SYN Flag Cnt \\
RST Flag Cnt & PSH Flag Cnt & ACK Flag Cnt \\
URG Flag Cnt & CWE Flag Count & ECE Flag Cnt \\
Down/Up Ratio & Pkt Size Avg & Fwd Seg Size Avg \\
Bwd Seg Size Avg & Fwd Byts/b Avg & Fwd Pkts/b Avg \\
Fwd Blk Rate Avg & Bwd Byts/b Avg & Bwd Pkts/b Avg \\
Bwd Blk Rate Avg & Subflow Fwd Pkts & Subflow Fwd Byts \\
Subflow Bwd Pkts & Subflow Bwd Byts & Init Fwd Win Byts \\
Init Bwd Win Byts & Fwd Act Data Pkts & Fwd Seg Size Min \\
Active Mean & Active Std & Active Max \\
Active Min & Idle Mean & Idle Std \\
Idle Max & Idle Min & \\
\hline
\end{tabular}
\label{table:cicflowmeter_features}
\end{table}

The CICFlowMeter has a command-line interface (which can be accessed by executing the Java main class \verb|cic.cs.unb.ca.ifm.Cmd|) and a graphical user interface (GUI, \verb|cic.cs.unb.ca.ifm.CICFlowMeter|). The command-line interface takes two parameters: the name of the input file (in \verb|pcap| format) or directory (in that case, all pcap files in that directory are read) and the name of the output directory. For each pcap file which was read, a separate comma-separated values (csv) file containing the extracted features is written to the output directory. The GUI allows to adjust additional parameters, such as the flow timeouts (see Figure \ref{fig:cicflowmeter_offline}), and is able to extract features in realtime by directly listening to a network interface (see Figure \ref{fig:cicflowmeter_realtime}).

\begin{figure}[htbp]  
\centerline{\includegraphics[scale=0.45]{CICFlowMeterOffline.png}}
\caption{CICFlowMeter GUI (offline mode)}
\label{fig:cicflowmeter_offline}
\end{figure}

\begin{figure}[htbp]  
\centerline{\includegraphics[scale=0.45]{CICFlowMeterRealtime.png}}
\caption{CICFlowMeter GUI (realtime mode)}
\label{fig:cicflowmeter_realtime}
\end{figure}

\subsection{Adaptation of the CICFlowMeter software}\label{adaptation}
Because the CICFlowMeter is open-source software available at GitHub \cite{cicflowmeter_github} and written in Java, it could easily be extended for the requirements of this paper.

The biggest change was necessary because of the way how the labelling was achieved. To ensure both accuracy and flexibility, the network traffic resulting from each host was recorded separately, resulting in multiple files in the pcap format \cite{paper1}. The label is defined by the file name: if the file name starts with an attack type label, that label is assigned to it - otherwise, its contents are labelled as benign network traffic. For this reason, the command-line interface of the CICFlowMeter was extended. It now retrieves a list of attack type labels, separated by a colon (e.g. \verb|ssh-bruteforce:bot:dos-goldeneye|) as an additional command-line parameter. It then reads the contents of the input directory specified in another command-line parameter and opens every file in it which name ends with \verb|.pcap|. So, instead of reading one file at a time as in the original implementation, multiple instances of the class \verb|cic.cs.unb.ca.jnetpcap.PacketReader| are created (one for each pcap file in the input directory), and assigned a label depending on the file name. Then, the information of the first valid packet is read from each \verb|PacketReader|. In the next step, the timestamps of these packets are compared, and the packet with the lowest timestamp (i.e. the earliest) is sent to a single \verb|FlowGenerator| instance. The only change made to the \verb|FlowGenerator| implementation is that it now takes the label as a parameter and forwards it to the \verb|BasicFlow| instance. The \verb|BasicFlow| keeps track of how many packets per label were found, and then sets the label of the flow accordingly: if all packets of that flow are benign, the label is set to \verb|BENIGN|, otherwise it is set to the attack type label with the highest packet count. After processing the packet data, the next packet from the same \verb|PacketReader| is read. Afterwards, the determination of the packet with the lowest timestamp is repeated, and its contents are also sent to the \verb|FlowGenerator|. When the end of file (EOF) has been reached for any pcap file, the corresponding \verb|PacketReader| is no longer considered, and the process continues with the remaining \verb|PacketReader| instances. This is repeated until all packets from all opened pcap files have been read.

Another change was made in order to efficiently use the newly created dataset with the neural network from \cite{max1}. For that reason, and to achieve as much flexibility as possible, another command-line parameter was added to the CICFlowMeter. This parameter lists all features which should be extracted (as with the attack labels, separated by a colon). The function which writes the flow features to the resulting \verb|csv| file was extended and now retrieves the list of flow features, which is passed from the newly added command-line parameter. Instead of writing all features supported by the CICFlowMeter, it now writes only the desired features, in the exact order they were specified. Of course, this also has to be considered when writing the headers to the first line of the \verb|csv| file (these list the names of the features). The new parameter allows the data to be extracted in a way that does not unnecessarily complicate the preprocessing of the neural network. If the parameter is ommitted, all features are written to the \verb|csv| file, exactly as in the original version of the CICFlowMeter.

For optimal integration into the infrastructure created in \cite{paper1}, a Docker image for feature extraction was created. For reasons of efficiency, the \verb|Dockerfile| used to create that image was written so that it employs a multi-stage build \cite{docker-multistage}. The first stage uses the image \verb|gradle:6.3.0-jdk11| (the Gradle and Java versions are set explicitly to ensure compatibility with the CICFlowMeter source code) from Docker Hub \cite{dockerhub}, copies the modified CICFlowMeter source code from the build context (in this case this is the directory that the \verb|Dockerfile| is contained in) into a temporary image and runs \verb|gradle| to build the CICFlowMeter distribution. The second stage is based on the image \verb|azul/zulu-openjdk:11| (again, the Java version is set explicitly). It runs \verb|jlink| \cite{jlink} to create a minimal Java 11 Runtime Environment containing only the modules required by the CICFlowMeter, which is written to another temporary image. Finally, the last stage builds the actual feature extraction Docker image. It is based on \verb|ubuntu:latest|, installs the \verb|libpcap| library required by the CICFlowMeter and copies the Java Runtime Environment and the CICFlowMeter distribution from the temporary images created in the previous stages. The \verb|ENTRYPOINT| of the docker image is set to the Java Virtual Machine binary, and all command-line parameters which are required to run the command-line interface of the CICFlowMeter are set as well. The data directory is set to \verb|/mnt/packet-data|, which means that the external directory containing all the pcap files has to be mounted to that directory of the Docker container. The \verb|csv| file with the extracted features is also written to that directory. Another dependency required by the Docker container is the environment variable \verb|ATTACKS|, which has to contain the attack labels separated by a colon. The list of features is read from the environment variable \verb|FEATURES|. When this environment variable is missing, all supported features are extracted.

\section{Neural Network}\label{nn}
The created dataset will be used in the further course of this paper to build a neural network. With this it should be possible to detect anomalies in network traffic (e. g. DoS attacks) but also harmless data packets. This is a continuation of  \cite{max1}, in which an attempt was made to create a neural network for intrusion detection on the basis of the publicly available CSE-CIC-IDS2018 dataset  \cite{max2}. The knowledge gained at that time is used, the source codes are reused accordingly and slightly adapted to the new dataset. In principle, the performance of neural networks over different datasets is to be checked. The following scenarios will be considered:
\smallskip

@@

\section{Conclusion}\label{conclusion}
In summary, the performance of neural networks in terms of intrusion detection over different datasets is rather mixed. If these machine learning algorithms are tested with the same dataset with which they were trained, the result is promising. Using another dataset for testing results in a poor classification performance. However, there is lots of potential for future research with neuronal networks in terms of intrusion detection. It seems useful to analyse in detail why the mentioned difficulties occur in the classification and which factors are relevant for the neural network. An intensive examination of the finding of optimal hyperparameters would also be desirable in the course of future work. Furthermore, it would also make sense to simplify the way classification is done. Instead, a simple classification like network traffic is “benign” or “not benign” would be conceivable. Moreover, it should be checked whether a better performance can be achieved with other machine learning algorithms.

\begin{thebibliography}{00}
\bibitem{paper1} B. Gally, S. Lipp and D. Marijanovic, "Automated Network Packet Generation for Evaluation of Neural Networks in Intrusion Prevention Systems", FH Technikum, Vienna, 30 June 2020.
\bibitem{max2} I. Sharafaldin, A. H. Lashkari and A. A. Ghorbani, "Toward Generating a New Intrusion Detection Dataset and Intrusion Traffic Characterization", in 4th International Conference on Information Systems Security and Privacy (ICISSP), Portugal, 2018.
\bibitem{max1} F. Ahrens, C. Cat and R. Freingruber, "A Neural-Network Approach for an Intrusion Detection System", FH Technikum, Vienna, 2019.
\bibitem{cicflowmeter} Canadian Institute for Cybersecurity (CIC), "Applications | Research | Canadian Institute for Cybersecurity | UNB", 2016. [Online]. Available: https://www.unb.ca/cic/research/applications.html\#CICFlowMeter. [Accessed 29 June 2020].
\bibitem{sopuru} J. Sopuru, A. Sari and M. Akkaya, "Modeling a malware detection and categorization system based on seven network flow-based features", in International Journal of Innovative Technology and Exploring Engineering (IJITEE), vol. 8(7), pp. 2982-2989, 05 2019
\bibitem{lashkari} A. H. Lashkari, G. D. Gil, M. S. I. Mamun and A. A. Ghorbani, "Characterization of tor traffic using time based features", in Proceedings of the 3rd International Conference on Information Systems Security and Privacy (ICISSP 2017), pp. 253-262, 2017.
\bibitem{wang} W. Wang, Y. Sheng, J. Wang, X. Zeng, X. Ye, Y. Huang and M. Zhu, HAST-IDS: Learning hierarchical spatial-temporal features Using deep neural networks to improve intrusion detection. IEEE Access, vol. 6, pp. 1792-1806, 2018.
\bibitem{moustafa} N. Moustafa and J. Slay, "UNSW-NB15: a comprehensive data set for network intrusion detection systems (UNSW-NB15 network data set)", in 2015 Military Communications and Information Systems Conference (MilCIS), Canberra, ACT, 2015, pp. 1-6.
\bibitem{celik} Z. Celik, Berkay, R. J. Walls, P. McDaniel and A. Swami, "Malware traffic detection using tamper resistant features", in MILCOM 2015 - 2015 IEEE Military Communications Conference, pp. 330–335, Oct 2015.
\bibitem{waizumi} Y. Waizumi, Y. Sato and Y. Nemoto, "A network-based anomaly detection system using multiple network features", in Proceedings of the Third International Conference on Web Information Systems and Technologies - Internet Technology, pp. 410-413, 2007.
\bibitem{cicflowmeter_github} GitHub - ahlashkari/CICFlowMeter: CICFlowmeter-V4.0 (formerly known as ISCXFlowMeter) is an Ethernet traffic Bi-flow generator and analyzer for anomaly detection that has been used in many Cybersecurity datsets such as Android Adware-General Malware dataset (CICAAGM2017), IPS/IDS dataset (CICIDS2017), Android Malware dataset (CICAndMal2017) and Distributed Denial of Service (CICDDoS2019). [Online]. Available: https://github.com/ahlashkari/CICFlowMeter. [Accessed 29 June 2020].
\bibitem{docker-multistage} Use multi-stage builds | Docker Documentation. [Online]. Available: https://docs.docker.com/develop/develop-images/multistage-build/. [Accessed 29 June 2020].
\bibitem{dockerhub} Docker Hub. [Online]. Available: https://hub.docker.com/. [Accessed 29 June 2020].
\bibitem{jlink} The Incredible Shrinking Java Platform - Azul Systems, Inc.. [Online]. Available: https://www.azul.com/blog/the-incredible-shrinking-java-platform/. [Accessed 29 June 2020].
\bibitem{max3} D. Meyer, "Performance Assessment \& Tuning - Presentation slides", FH Technikum, Vienna, 2018.
\bibitem{max4} J. Brownlee, "How to Calculate Correlation Between Variables in Python", 3 June 2020. [Online]. Available: https://machinelearningmastery.com/how-to-use-correlation-to-understand-the-relationship-between-variables/. [Accessed 27 June 2020].
\bibitem{max5} S. Norena, "Finding Correlation Between Many Variables (Multidimensional Dataset) with Python", 26 April 2018. [Online]. Available: https://medium.com/@sebastiannorena/finding-correlation-between-many-variables-multidimensional-dataset-with-python-5deb3f39ffb3. [Accessed 27 June 2020].
\end{thebibliography}
